MATH 245 old and unused content

### 2024-02-13 PDEs/math245_hw2.Rmd
more PDEs

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-12"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

@@

@@

@@

where $g(x)=-g(x)$ is an even function. Hint: set $w=xu$.

## Problem 8   
Solve the following characteristic initial value problem:
\begin{equation} 
\begin{cases}
  y^3u_{xx}-yu_{yy}+u_y=0 & 0<x<4, \quad \lvert y \rvert \leq 2 \sqrt{2} \\
  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{for} 2 \leq x \leq 4 \\
  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{for} 0 \leq x \leq 2 \\
\end{cases}
\end{equation}

>>>>

---
title: "MATH 245 Homework 2"
author: "Ruby Krasnow and Tommy Thach"
date: "2024-02-13"
output:
  pdf_document:
    extra_dependencies: ["derivative", "tcolorbox", "xcolor"]
---

@@

## Done, feel OK: 1, 2
## Done, very low confidence: 4, 6
## Mostly done: 7
## Not started: 3, 5, 8

@@

Let's try the factorization method if $c=c(x)$:
$$\left( \partial_{t}+c(x)\partial_{x}\right) \left(\partial_{t}-c(x)\partial_{x}\right)u=0$$
Set $\xi=x+c(x)t$ and $\eta=x-c(x)t$.

By the chain rule,

$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}$$

$$=\partial_{\xi}(1+c')+\partial_{\eta}\,(1-c')$$

$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}$$
$$=c\partial_{\xi}-c\partial_{\eta}$$
When we plug these back into
$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
we get

$$ \left(c\partial_{\xi}-c\partial_{\eta}+c\partial_{\xi}(1+c')+c\partial_{\eta}\,(1-c')\right)
\left(c\partial_{\xi}-c\partial_{\eta} -c\partial_{\xi}(1+c')-c\partial_{\eta}\,(1-c')\right)u=0$$

When $c$ is constant, $c'=0$ and this simplifies to $-4c^2 u_{\xi\eta}=0$, allowing us to integrate to find $u(x,t)=F(x+ct)+G(x-ct)$. But we cannot do the same simplification when we have the $c'$ terms, meaning we cannot solve the wave equation by the same method when the wave speed is not constant.

Similarly,

Let's try the factorization method if $c=c(t)$:
$$\left( \partial_{t}+c(t)\partial_{x}\right) \left(\partial_{t}-c(t)\partial_{x}\right)u=0$$
Set $\xi=x+c(t)t$ and $\eta=x-c(t)t$.

By the chain rule,

$$\frac{\partial}{\partial x} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial x}+\frac{\partial}{\partial\eta}\frac{\partial\eta}{\partial x}=
\partial_{\xi}+\partial_{\eta}$$

$$\frac{\partial}{\partial t} = \frac{\partial}{\partial \xi} \frac{\partial \xi}{\partial t}+\frac{\partial }{\partial\eta}\frac{\partial\eta}{\partial t}=\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)
$$

When we plug these back into
$$\left( \partial_{t}+c\partial_{x}\right) \left(\partial_{t}-c\partial_{x}\right)u=0$$
we get

$$\left( \partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)+c\partial_{\xi}+c\partial_{\eta} \right)
\left(\partial_{\xi}(c't+c)-\partial_{\eta}(c't+c)-c\partial_{\xi}-c\partial_{\eta}\right)u=0$$
Where again we cannot cancel terms to simplify our equation as we could with constant $c$.

@@

where $g(x)=g(-x)$ is an even function. Hint: set $w=xu$.

Using the results from when we set $w=ru$ in Problem 4,
$$u_t=\frac{w_t}{x}, u_{tt}=\frac{w_{tt}}{x} $$
$$u_x= \frac{w_x}{x}- \frac{w}{x^2} $$

$$u_{xx}=  \frac{w_{xx}}{x}-\frac{2w_x}{x^2}+\frac{2w}{x^3}$$
So $u_{tt}-u_{xx}- \frac{2}{x}u_x=0$ becomes

$$\frac{w_{tt}}{x} -\frac{w_{xx}}{x}+\frac{2w_x}{x^2}- \frac{2w}{x^3}- \frac{2}{x} \left(\frac{w_x}{x}- \frac{w}{x^2}\right)=0 $$
Which simplifies to $w_{tt}-w_{xx}=0$, the wave equation with $c=1$.
By the D'Alembert formula, since we have $\varphi=0$,
$$w(t,x)=\frac{1}{2c} \int_{x-t}^{x+t}{\psi(s)ds}$$
$w_t=xu_t$
Since $y=x$ is an odd function and $u_t(0,x)=g(x)$ is an even function, $w_t=xg(x)=h(x)$ must be an odd function.

## Problem 8   
Solve the following characteristic initial value problem:
\begin{equation} 
\begin{cases}
  y^3u_{xx}-yu_{yy}+u_y=0 & 0<x<4, \quad \lvert y \rvert \leq 2 \sqrt{2} \\
  u(x,y)= f(x) & x+ \frac{y^2}{2}=4 \text{ for } 2 \leq x \leq 4 \\
  u(x,y)=g(x) & x- \frac{y^2}{2}=0 \text{ for } 0 \leq x \leq 2 \\
\end{cases}
\end{equation}

### 2024-02-13 PDEs/math245_old.Rmd
math hw proiblem 2
math hw2 num 4
more PDEs

>>>>

Then set $\xi=x+3y, \; \eta=x-y$, giving us

$$\partial_x= \pdv{}{\xi} \pdv{\xi}{x}+\pdv{}{\eta} \pdv{\eta}{x}=\partial_{\xi}+\partial_{\eta}$$
$$\partial_y= \pdv{}{\xi} \pdv{\xi}{y}+\pdv{}{\eta} \pdv{\eta}{y}= 3\partial_{\xi}-\partial_{\eta}$$
Which means that 
$$\partial_{x}+3 \partial_{y}= 10 \partial_{\xi}-2\partial_{\eta}$$

$$\partial_{x}-\partial_{y}=-2\partial_{\xi}+2\partial_{\eta} $$

Which means we can rewrite the original equation as 
$$\left( \partial_{x}+3\partial_{y}\right) \left(\partial_{x}-\partial_{y}\right)u=0 \quad\longrightarrow\quad -8\partial$$
\begin{equation}\label{eq:sphere1}
x=r \sin{\theta} \cos{\phi}, \quad y=r \sin{\theta} \sin{\phi}, \quad z=r \cos{\theta}
\end{equation}

\begin{equation}\label{eq:sphere2}
r=\sqrt{x^2+y^2+z^2}, \quad \theta=\arccos{(\frac{z}{r})}, \quad \phi=\arctan{(\frac{y}{x})} \end{equation}

$$\theta=\arccos{(\frac{z}{r})}$$

Using the chain rule:
$$\Delta u= \frac{{\partial}^{2}f}{\partial x^{2}}+\frac{{\partial}^{2}f}{\partial y^{2}}+\frac{{\partial}^{2}f}{\partial z^{2}}$$

Now we find the following values from (\ref{eq:sphere1}):
$$\frac{\partial r}{\partial x}=\sin(\theta)\cos(\phi),
\qquad\frac{\partial\theta}{\partial x}=\frac{1}{r}\cos(\theta)\cos(\phi),
\qquad\frac{\partial\phi}{\partial x}=-\frac{1}{r}\frac{\sin(\phi)}{\sin(\theta)}$$
$$\frac{\partial r}{\partial y}=\sin(\theta)\sin(\phi),\qquad\frac{ \partial\theta}{\partial y}=\frac{1}{r}\cos(\theta)\sin(\phi),\qquad\frac{\partial\phi}{\partial y}=\frac{1}{r}\frac{\cos(\phi)}{\sin(\theta)} $$

$$\frac{\partial r}{\partial z}=\cos(\theta),\qquad\frac{\partial \theta}{\partial z}=-\frac{1}{r}\sin(\theta),\qquad\frac{\partial\phi}{\partial z}=0$$
$$ \frac{\partial u}{\partial x} = \frac{\partial u}{\partial r} \frac{\partial r}{\partial x}+\frac{\partial u}{\partial\theta}\frac{\partial\theta}{\partial x}+\frac{\partial u}
{\partial\phi}\frac{\partial\phi}{\partial x}=u_r\sin(\theta)\cos(\phi)$$

$$\frac{\partial u_x}{\partial x} = \frac{\partial u_x}{\partial r} \frac{\partial r}{\partial x}+\frac{\partial u_x}{\partial\theta}\frac{\partial\theta}{\partial x}+\frac{\partial u_x}{\partial\phi}\frac{\partial\phi}{\partial x}= $$

$$u_{rr}\sin^2(\theta)\cos^2(\phi)+ \frac{u_r}{r} \cos^2(\theta)\cos^2(\phi)- $$

$$u(x,t)= \frac{\varphi(x+ct)+\varphi(x-ct)}{2}+\frac{1}{2c} \int_{x-ct}^{x+ct}{\psi(s)ds}$$
Assume $c$ is not a constant, but instead a function of $x$.
